In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-05 paper/supplementary_main.tex
Updates from Overleaf

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general,
abstract form here, and to the reference~\cite{masri2023modelling} for
the derivation and analysis of this 3D-1D model. Recall that we model
a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here. We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
  %= f
   = 0
  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivacular networks $\Lambda_a, \Lambda_v$. To summarize,
\begin{equation*}
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\end{equation*}
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

@@

>>>>

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general,
abstract form here, and to the reference~\cite{masri2024modelling} for
the derivation and analysis of this 3D-1D model. Recall that we model
a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here. We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
  %= f
   = 0
  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivacular networks $\Lambda_a, \Lambda_v$. To summarize,
\begin{equation*}
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\end{equation*}
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{paper/figures/mesh_refinement.png}
    \caption{Illustration of the three different meshes: low resolution, standard resolution, high resolution (from left to right).}
    \label{fig:mesh_refinement}
\end{figure}
